---
title: "Influencial points checking"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages + options, include=FALSE}
if (!require("pacman")){
  install.packages("pacman", repos='http://cran.us.r-project.org')
}

p_load("tidyverse", "magrittr", "plyr", "haven", "here", "broom",
       "survival", "survminer")

#No scientific notation
options(scipen = 999)

set.seed(20200819)
```

```{r directories + source scripts, include=FALSE}
#---- note ----
# Since the difference between win and OS, put substituted directory here
# Yingyan's directory: C:/Users/yingyan_wu
#                      C:/Users/yingyan_wu/Dropbox
# Crystal's directory: /Users/CrystalShaw
#                     ~/Dropbox/Projects
# MRG desktop directory: C:/Users/cshaw/Dropbox/Projects

#Changing directories here will change them throughout the script
path_to_dropbox <- "C:/Users/yingyan_wu/Dropbox"

#---- source scripts ----
# source(here::here("RScripts", "mask.R"))
# source(here::here("RScripts", "mask_impute_pool.R"))
```

```{r load the data}
CESD_data_wide <- 
  read_csv(paste0(path_to_dropbox, 
                  "/exposure_trajectories/data/", 
                  "CESD_data_wide.csv"), 
           col_types = cols(HHIDPN = col_character())) 

for(wave in seq(4, 9)){
  CESD_data_wide %<>% 
    mutate(!!paste0("r", wave, "cesd_death2018") := 
             !!sym(paste0("r", wave, "cesd"))*death2018, 
           !!paste0("r", wave - 1, "cesd_conde_impute") := 
             !!sym(paste0("r", wave - 1, "cesd"))*
             !!sym(paste0("r", wave - 1, "conde_impute")))
}
```

## Influence points dx
Large magnitude of the score residual of an individual with respect to a particular covariate indicates heavy influence of that individual in the estimation of the regression effect of that covariate

Large magnitude of the deviance residual for an observation indicates that it is a potential outlier to the model.

### CESD wave 4
```{r CESD wave 4}
#---- CES-D Wave 4 ----
TTEmodel_CESD4 <- 
  coxph(Surv(survtime, observed) ~ r4not_married_partnered + r4widowed + 
          ed_cat + r4drinking_cat + r4memrye_impute + r4stroke_impute + 
          r4hearte_impute + r4lunge_impute + r4cancre_impute + r4hibpe_impute + 
          r4diabe_impute + smoker + r4BMI + hispanic + black + other + female + 
          r4age_y_int + r4cesd_elevated, data = CESD_data_wide)
# summary(TTEmodel_CESD4)
tidy(TTEmodel_CESD4, exponentiate = FALSE, conf.int = TRUE)

#---- High-leverage points ----
# Can't run in a coxph setting
# highleverage <- function(fit) {
#   p <- length(coefficients(fit))
#   n <- length(fitted(fit))
#   ratio <- p/n
#   plot(hatvalues(fit), main = "Index Plot of Ratio")
#   abline(h=c(2,3)*ratio, col="red",  lty=2)
#   identify(1:n, hatvalues(fit), names(hatvalues(fit)))
# }
# highleverage(TTEmodel_CESD4)

#---- ** dfbeta ----
(threshold <- 2/sqrt(nrow(CESD_data_wide)))
# An attempt to visualize the dfbetas
# dfbeta function is not for coxph settings
# dfbetas1 <- as.data.frame(dfbetas(TTEmodel_CESD4))

dfbeta_residual_cesd4 <- as.data.frame(resid(TTEmodel_CESD4, type = "dfbeta"))
colnames(dfbeta_residual_cesd4) <- names(stats::coef(TTEmodel_CESD4))
dfbeta_residual_cesd4 %<>% cbind(CESD_data_wide %>% select(HHIDPN)) %>%
  mutate(obsid = 1:nrow(dfbeta_residual_cesd4)) %>%
  select(obsid, HHIDPN, everything())

# Plot
ggcoxdiagnostics(TTEmodel_CESD4, type = "dfbeta",
                 linear.predictions = FALSE, ggtheme = theme_bw(),
                 title = "DFBETA residuals plot (E: CESD wave 4)")
  # geom_hline(aes(yintercept = threshold), color = "red") +
  # geom_hline(aes(yintercept = -threshold), color = "red")
ggsave(paste0(path_to_dropbox, "/exposure_trajectories/",
              "manuscript/figures/influential_points_dx/", 
              "dfbeta_cesd4.jpeg"), 
       device = "jpeg", dpi = 300, width = 11, height = 9, units = "in")

#---- Deviance residuals ----
ggcoxdiagnostics(TTEmodel_CESD4, type = "deviance",
                linear.predictions = FALSE, ggtheme = theme_bw(),
                title = "Deviance residuals plot (E: CESD wave 4)")
ggsave(paste0(path_to_dropbox, "/exposure_trajectories/",
              "manuscript/figures/influential_points_dx/", 
              "deviance_cesd4.jpeg"), 
       device = "jpeg", dpi = 300, width = 11, height = 9, units = "in")

#---- **Score residuals----
ggcoxdiagnostics(TTEmodel_CESD4, type = "score",
                 linear.predictions = FALSE, ggtheme = theme_bw(),
                 title = "score residuals plot (E: CESD wave 4)")
ggsave(paste0(path_to_dropbox, "/exposure_trajectories/",
              "manuscript/figures/influential_points_dx/", 
              "deviance_cesd4.jpeg"), 
       device = "jpeg", dpi = 300, width = 11, height = 9, units = "in")

```
### CES-D Wave 9
```{r CES-D wave 9}
TTEmodel_CESD9 <- 
  coxph(Surv(survtime, observed) ~ r9not_married_partnered + r9widowed + 
          ed_cat + r9drinking_cat + r9memrye_impute + r9stroke_impute + 
          r9hearte_impute + r9lunge_impute + r9cancre_impute + r9hibpe_impute + 
          r9diabe_impute + smoker + r9BMI + hispanic + black + other + female + 
          r9age_y_int + r9cesd_elevated, data = CESD_data_wide)
# summary(TTEmodel_CESD9)
tidy(TTEmodel_CESD9, exponentiate = FALSE, conf.int = TRUE)

#---- **dfbeta ----
dfbeta_residual_cesd9 <- as.data.frame(resid(TTEmodel_CESD9, type = "dfbeta"))
colnames(dfbeta_residual_cesd9) <- names(stats::coef(TTEmodel_CESD9))
dfbeta_residual_cesd9 %<>% cbind(CESD_data_wide %>% select(HHIDPN)) %>%
  mutate(obsid = 1:nrow(dfbeta_residual_cesd9)) %>%
  select(obsid, HHIDPN, everything())

# Plots
ggcoxdiagnostics(TTEmodel_CESD9, type = "dfbeta",
                 linear.predictions = FALSE, ggtheme = theme_bw(),
                 title = "DFBETA residuals plot (E: CESD wave 9)")
ggsave(paste0(path_to_dropbox, "/exposure_trajectories/",
              "manuscript/figures/influential_points_dx/", 
              "dfbeta_cesd9.jpeg"), 
       device = "jpeg", dpi = 300, width = 11, height = 9, units = "in")

ggcoxdiagnostics(TTEmodel_CESD9, type = "deviance",
                linear.predictions = FALSE, ggtheme = theme_bw(),
                title = "Deviance residuals plot (E: CESD wave 9)")
ggsave(paste0(path_to_dropbox, "/exposure_trajectories/",
              "manuscript/figures/influential_points_dx/", 
              "deviance_cesd9.jpeg"), 
       device = "jpeg", dpi = 300, width = 11, height = 9, units = "in")

# Score?

```

### Elevated average CESD
```{r elevated average CESD}
TTEmodel_elevated_avg_CESD <- 
  coxph(Surv(survtime, observed) ~ r4not_married_partnered + r4widowed + 
          ed_cat + r4drinking_cat + r4memrye_impute + r4stroke_impute + 
          r4hearte_impute + r4lunge_impute + r4cancre_impute + r4hibpe_impute + 
          r4diabe_impute + smoker + r4BMI + hispanic + black + other + female + 
          r4age_y_int + avg_cesd_elevated, data = CESD_data_wide)
# summary(TTEmodel_elevated_avg_CESD)
tidy(TTEmodel_elevated_avg_CESD, exponentiate = FALSE, conf.int = TRUE)

#---- **dfbeta ----
dfbeta_residual_avg_cesd <- as.data.frame(resid(TTEmodel_elevated_avg_CESD, 
                                                type = "dfbeta"))
colnames(dfbeta_residual_avg_cesd) <- 
  names(stats::coef(TTEmodel_elevated_avg_CESD))
dfbeta_residual_avg_cesd %<>% cbind(CESD_data_wide %>% select(HHIDPN)) %>%
  mutate(obsid = 1:nrow(dfbeta_residual_avg_cesd)) %>%
  select(obsid, HHIDPN, everything())

# plots
ggcoxdiagnostics(TTEmodel_elevated_avg_CESD, type = "dfbeta",
                 linear.predictions = FALSE, ggtheme = theme_bw(),
                 title = "DFBETA residuals plot (E: Elevated avg CESD)")
ggsave(paste0(path_to_dropbox, "/exposure_trajectories/",
              "manuscript/figures/influential_points_dx/", 
              "dfbeta_elevated_avg_cesd.jpeg"), 
       device = "jpeg", dpi = 300, width = 11, height = 9, units = "in")

ggcoxdiagnostics(TTEmodel_elevated_avg_CESD, type = "deviance",
                 linear.predictions = FALSE, ggtheme = theme_bw(),
                 title = "Deviance residuals plot (E: Elevated avg CESD)")
ggsave(paste0(path_to_dropbox, "/exposure_trajectories/",
              "manuscript/figures/influential_points_dx/", 
              "deviance_elevated_avg_cesd.jpeg"), 
       device = "jpeg", dpi = 300, width = 11, height = 9, units = "in")
```

### Elevated CES-D count
```{r Total count of elevated CES-D}
TTEmodel_total_CESD <- 
  coxph(Surv(survtime, observed) ~ r4not_married_partnered + r4widowed + 
          ed_cat + r4drinking_cat + r4memrye_impute + r4stroke_impute + 
          r4hearte_impute + r4lunge_impute + r4cancre_impute + r4hibpe_impute + 
          r4diabe_impute + smoker + r4BMI + hispanic + black + other + female + 
          r4age_y_int + total_elevated_cesd, data = CESD_data_wide)
# summary(TTEmodel_total_CESD)
tidy(TTEmodel_total_CESD, exponentiate = FALSE, conf.int = TRUE)

#---- ** dfbeta ----
dfbeta_residual_total_cesd <- as.data.frame(resid(TTEmodel_total_CESD, 
                                                  type = "dfbeta"))
colnames(dfbeta_residual_total_cesd) <- names(stats::coef(TTEmodel_total_CESD))
dfbeta_residual_total_cesd %<>% cbind(CESD_data_wide %>% select(HHIDPN)) %>%
  mutate(obsid = 1:nrow(dfbeta_residual_total_cesd)) %>%
  select(obsid, HHIDPN, everything())

ggcoxdiagnostics(TTEmodel_total_CESD, type = "dfbeta",
                 linear.predictions = FALSE, ggtheme = theme_bw(),
                 title = "DFBETA residuals plot (E: Elevated CESD count)")
ggsave(paste0(path_to_dropbox, "/exposure_trajectories/",
              "manuscript/figures/influential_points_dx/", 
              "dfbeta_cesd_total_cesd.jpeg"), 
       device = "jpeg", dpi = 300, width = 11, height = 9, units = "in")

ggcoxdiagnostics(TTEmodel_total_CESD, type = "deviance",
                 linear.predictions = FALSE, ggtheme = theme_bw(),
                 title = "Deviance residuals plot (E: Elevated CESD count)")
ggsave(paste0(path_to_dropbox, "/exposure_trajectories/",
              "manuscript/figures/influential_points_dx/", 
              "deviance_total_cesd.jpeg"), 
       device = "jpeg", dpi = 300, width = 11, height = 9, units = "in")
```

### Influential points
```{r influential points}
# CES-D wave 4
cesd4_point <- dfbeta_residual_cesd4 %>% 
  filter((obsid < 2500 | obsid > 5000) & r4cesd_elevated < -0.01) %>%
  print()
# CES-D wave 9
cesd9_point <- dfbeta_residual_cesd9 %>% 
  filter(obsid < 2500 & r9cesd_elevated < -0.015) %>%
  print()

# CES-D elevated avg CES-D
cesd_elevated_avg <- dfbeta_residual_avg_cesd %>% 
  filter((obsid < 2500 | obsid > 7500) & avg_cesd_elevated < -0.01) %>%
  print()

influential_df <- 
  full_join(cesd4_point, cesd9_point, by = c("obsid", "HHIDPN")) %>%
  full_join(., cesd_elevated_avg, by = c("obsid", "HHIDPN"))
write_csv(influential_df, paste0(path_to_dropbox,
                           "/exposure_trajectories/",
                           "inducing_missingness_troubleshooting/",
                           "model_method/",
                           "influential_df.csv"))
```

### Cox ph assumptions
```{r}

(testph1 <- cox.zph(TTEmodel_CESD4, transform = "km", global = T)) 
# Global P = 0.0016, violate to the assumption of proportional hazards
ggcoxzph(testph1)

(testph2 <- cox.zph(TTEmodel_CESD9, transform = "km", global = T)) 
# Global P = 0.00000000055, violate to the assumption of proportional hazards
ggcoxzph(testph2)

(testph3 <- cox.zph(TTEmodel_elevated_avg_CESD, transform = "km", global = T)) 
# Global P = 0.0022, violate to the assumption of proportional hazards
ggcoxzph(testph3)

(testph4 <- cox.zph(TTEmodel_total_CESD, transform = "km", global = T)) 
# Global P = 0.0016, violate to the assumption of proportional hazards
ggcoxzph(testph4)

# KM curve
ph_cesd4 <- surv_fit(Surv(survtime, observed) ~ 
                    r4cesd_elevated, data = CESD_data_wide)
ggsurvplot(ph_cesd4, 
           data = CESD_data_wide)

ph_cesd_9 <- surv_fit(Surv(survtime, observed) ~ 
                       r9cesd_elevated, data = CESD_data_wide)
ggsurvplot(ph_cesd_9, 
           data = CESD_data_wide)

ph_elevated_avg_cesd <- surv_fit(Surv(survtime, observed) ~ 
                                   avg_cesd_elevated, data = CESD_data_wide)
ggsurvplot(ph_elevated_avg_cesd, 
           data = CESD_data_wide)

ph_total_elevated_cesd <- surv_fit(Surv(survtime, observed) ~ 
                                    total_elevated_cesd, data = CESD_data_wide)
ggsurvplot(ph_total_elevated_cesd, 
           data = CESD_data_wide)
# But the KM curve showed that the PH assumptions were not violated???
```
### Influence plots
```{r}

```

