---
title: "Exposure Trajectories"
subtitle: "Data Subsets: CESD --> Mortality"
date: "`r format(Sys.time(), '%m/%d/%y')`"
output: 
  html_document:
    theme: paper
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages}
if (!require("pacman")){
  install.packages("pacman", repos='http://cran.us.r-project.org')
}

p_load("here", "tidyverse", "magrittr", "broom", "kableExtra", "survival")

#No scientific notation
options(scipen = 9)
```

```{r import data}
#Changing directories here will change them throughout the script
# Yingyan's directory: C:/Users/yingyan_wu
#                      C:/Users/yingyan_wu/Dropbox
# Crystal's directory: /Users/CrystalShaw
#                     ~/Dropbox/Projects
path_to_dropbox <- "~/Dropbox/Projects"

hrs_samp <- read_csv(paste0(path_to_dropbox,
                            "/exposure_trajectories/data/",
                            "hrs_samp_6CESD_waves4-9.csv")) 
```

```{r create new variables}
hrs_samp %<>% 
  mutate("med_CESD" = hrs_samp %>% 
           dplyr::select(paste0("r", seq(4, 9, by = 1), "cesd")) %>% 
           apply(., 1, function(x) median(x))) %>% 
  mutate("elevated_depress_sx_med" = ifelse(med_CESD > 4, 1, 0), 
         "elevated_depress_sx_last" = ifelse(r9cesd > 4, 1, 0))

# #Sanity check
# View(hrs_samp %>%
#        dplyr::select(paste0("r", seq(4, 9, by = 1), "cesd"), "med_CESD", 
#                      "elevated_depress_sx"))
# 
# summary(hrs_samp$med_CESD)
# table(hrs_samp$elevated_depress_sx_med, useNA = "ifany")
# table(hrs_samp$elevated_depress_sx_last, useNA = "ifany")
```

## Age Data

Distributions of age at last CESD Measure overlayed with age at death. I think we decided last time that we would use study time as our timescale so the overlap matters less?  

```{r age distributions}
plot_data <- hrs_samp %>% 
  dplyr::select("4age_y_int")

ggplot(data = plot_data) + 
  geom_bar(aes(x = `4age_y_int`, fill = "First CESD Wave"), stat = "count", 
           color = "lightblue", fill = "lightblue") + 
  theme_minimal() + xlab("Age") + 
  ggtitle("Age at first CESD measure")

plot_data <- hrs_samp %>% 
  dplyr::select("9age_y_int", "death2018", "age_death_y")

ggplot(data = plot_data) + 
  geom_bar(aes(x = `9age_y_int`, fill = "Last CESD Wave"), stat = "count", 
           color = "lightblue", fill = "lightblue") + 
  geom_bar(aes(x = age_death_y, fill = "Death by 2018"), stat = "count", 
           alpha = 0.5) +
  theme_minimal() + xlab("Age") + 
  ggtitle("Age at death overlayed on Age at last CESD measure") +
  # the labels must match what you specified above
  scale_fill_manual(name = "", 
                    values = c("Last CESD Wave" = "lightblue", 
                               "Death by 2018" = "gray"))
```

## Risk Regressions
Sample size: n = `r nrow(hrs_samp)`  
Proportion dead by 2018: `r round(mean(hrs_samp$death2018), 2)`  

Look at whether models for median CES-D give us tight enough CI  

This model categorizes elevated depressive symptoms based on median of all 6 CESD measures

```{r RRmodel1: median CESD}
RRmodel1_elevated_depress_sx_med <- 
  glm(death2018 ~ `9age_y` + female + hispanic + black + other + 
          elevated_depress_sx_med, family = poisson(link = "log"), 
      data = hrs_samp)
  
kable(tidy(RRmodel1_elevated_depress_sx_med, exponentiate = TRUE, 
           conf.int = TRUE)) %>% kableExtra::kable_styling()

RRmodel2_elevated_depress_sx_med <- 
  glm(death2018 ~ `9age_y` + female + hispanic + black + other + 
          elevated_depress_sx_med + smoker + as.factor(drinking9_cat) + as.factor(r9mstat_cat)
        , family = poisson(link = "log"), 
      data = hrs_samp)
  
kable(tidy(RRmodel2_elevated_depress_sx_med, exponentiate = TRUE, 
           conf.int = TRUE)) %>% kableExtra::kable_styling()
```

This model categorizes elevated depressive symptoms based on last CESD measure  

```{r RRmodel1: last CESD}
RRmodel1_elevated_depress_sx_last <- 
  glm(death2018 ~ `9age_y` + female + hispanic + black + other + 
          elevated_depress_sx_last, family = poisson(link = "log"), 
      data = hrs_samp)
  
kable(tidy(RRmodel1_elevated_depress_sx_last, exponentiate = TRUE, 
           conf.int = TRUE)) %>% kableExtra::kable_styling()

RRmodel2_elevated_depress_sx_last <- 
  glm(death2018 ~ `9age_y` + female + hispanic + black + other + 
          elevated_depress_sx_last + smoker + as.factor(drinking9_cat) + as.factor(r9mstat_cat)
        , family = poisson(link = "log"), 
      data = hrs_samp)
  
kable(tidy(RRmodel2_elevated_depress_sx_last, exponentiate = TRUE, 
           conf.int = TRUE)) %>% kableExtra::kable_styling()
```

```{r RRmodel2: med CESD}
# #@YW: please add other variables to this model based on the lit review and 
# #     change to be consistent with med CESD model above), then we can 
# #     uncomment this code
# model2_avg_BMI_cat <- 
#   glm(death2018 ~ `9age_y` + female + hispanic + black + other + 
#           as.factor(avg_BMI_cat) + smoker + as.factor(drinking9_cat) + 
#         as.factor(r9mstat_cat), family = poisson(link = "log"), 
#       data = hrs_samp)
#   
# kable(tidy(model2_avg_BMI_cat, exponentiate = TRUE, conf.int = TRUE)) %>% 
#   kableExtra::kable_styling()
```

```{r RRmodel2: last CESD}
# #@YW: please add other variables to this model based on the lit review and 
# #     change to be consistent with last CESD model above), then we can 
# #     uncomment this code
# model2_avg_BMI_cat <- 
#   glm(death2018 ~ `9age_y` + female + hispanic + black + other + 
#           as.factor(avg_BMI_cat) + smoker + as.factor(drinking9_cat) + 
#         as.factor(r9mstat_cat), family = poisson(link = "log"), 
#       data = hrs_samp)
#   
# kable(tidy(model2_avg_BMI_cat, exponentiate = TRUE, conf.int = TRUE)) %>% 
#   kableExtra::kable_styling()
```

## Time-to-Event
These analyses are more in line with what researchers would actually conduct. I did a really rough approximation of survival time by using the difference between a participant's age at last interview and their age at death. We can use actual dates in the real analysis if we want.

```{r format survival data}
hrs_samp %<>% mutate(survtime = age_death_y - `9age_y_int`) %>% 
  mutate(survtime = ifelse(is.na(survtime), `9age_y_int` + 10, survtime), 
         observed = ifelse(is.na(age_death_y), 0, 1))

# #Sanity check
# View(hrs_samp %>% dplyr::select("age_death_y", "9age_y_int", "survtime", 
#                                 "observed"))
```

This model categorizes elevated depressive symptoms based on median CESD measure  

```{r TTEmodel1: median CESD}
TTEmodel1_elevated_depress_sx_med <- 
  coxph(Surv(survtime, observed) ~ `9age_y` + female + hispanic + black + 
          other + elevated_depress_sx_med, data = hrs_samp)
  
kable(tidy(TTEmodel1_elevated_depress_sx_med, exponentiate = TRUE, 
           conf.int = TRUE)) %>% kableExtra::kable_styling()

TTEmodel2_elevated_depress_sx_med <- 
  coxph(Surv(survtime, observed) ~ `9age_y` + female + hispanic + black + 
          other + elevated_depress_sx_med
        + smoker + as.factor(drinking9_cat) + as.factor(r9mstat_cat), data = hrs_samp)
  
kable(tidy(TTEmodel2_elevated_depress_sx_med, exponentiate = TRUE, 
           conf.int = TRUE)) %>% kableExtra::kable_styling()
```

This model categorizes elevated depressive symptoms based on last measure

```{r TTEmodel1: last CESD}
TTEmodel1_elevated_depress_sx_last <- 
  coxph(Surv(survtime, observed) ~ `9age_y` + female + hispanic + black + 
          other + elevated_depress_sx_last, data = hrs_samp)
  
kable(tidy(TTEmodel1_elevated_depress_sx_last, exponentiate = TRUE, 
           conf.int = TRUE)) %>% kableExtra::kable_styling()

TTEmodel2_elevated_depress_sx_last <- 
  coxph(Surv(survtime, observed) ~ `9age_y` + female + hispanic + black + 
          other + elevated_depress_sx_last + 
        smoker + as.factor(drinking9_cat) + as.factor(r9mstat_cat), data = hrs_samp)
  
kable(tidy(TTEmodel2_elevated_depress_sx_last, exponentiate = TRUE, 
           conf.int = TRUE)) %>% kableExtra::kable_styling()
```
